On Sample Size for Block Bootstrap Confidence Intervals to Have Desired Coverage Rates

### 2023-08-11 Manuscript/manuscript.tex
Addressed Some Comments Still Need to Look over Everything

The necessary sample size for bootstrap standard errors to provide valid
uncertainty measures in practice has not been extensively studied. For
independent data and non-block bootstrap, \citet{hesterberg2015teachers} notes
that while percentile-based CIs \eds{for the mean? differences in means?} from 
nonparametric bootstrap
are more accurate than $t$-intervals for larger sample sizes, they are
less accurate for smaller sample sizes. \citet{chernick2009revisiting} finds 
that the best choice of method for estimating
the parameter of a distribution is dependent on the sample size, the number of
bootstrap resamples, and the confidence level. In the context of structural equation
modeling, \citet{nevitt2001performance} find
that a sample size of 200--1000 is usually sufficient for interval estimation
using standard nonparametric bootstrap. For dependent data, the necessary
sample
size for bootstrap is even less clear. In the setting of linear regression with a
focus on estimating the proportion of 
variation attributable to a certain variable, \citet{burch2012nonparametric} finds
that three nonparametric bootstrap methods (Standard, Percentile, and Bias-Corrected
and
Accelerated) approach the coverage of a pivotal quantity as $n$ gets larger for an
underlying normal distribution. However, they find that for other distributions, 
coverage may deteriorate as $n$ increases. 
When attempting to create interval estimations for the Pearson correlation 
coefficient of bivariate normal data, \citet{puth2015variety} find bootstrap 
methods to be less accurate than the Fisher's transformation method even 
for a sample of size $n = 100$ and correlation coefficient equivalent to 0.
Few studies have provided a practical guide on what size is needed
for block bootstrap inference for dependent data.
In a linear regression setting
with
dependent data where the
regression errors are generated from a homoscedastic AR(1) process, 
\citet{goncalves2005bootstrap} find that standard error
estimates from moving 
block bootstrap in small samples may be more accurate than
inference from closed-form asymptotic estimates, but moving block bootstrap 
\jy{did they do block or standard bootstrap?}
\mc{moving block bootstrap}
CIs still do not have sufficient coverage of the target parameter even for sample size
$n = 1024$.
\jy{I added more references to the reference folder that studied coverage of
  bootstrap CIs but not block bootstrap CIs.}

@@

\paragraph{Standard Normal CI}
Assuming that $\hat\theta_n$ is asymptotically normally distributed with
$\theta$ as the mean, we just need an estimate of the standard error to
construct an approximate CI \citep[p.168]{efron1993introduction}.
Let $\widehat{\text{SE}}$ be the empirical standard error of the block bootstrap
copies of $\hat\theta_n^{(b)}$, $b = 1, \ldots, B$.
A $(1 - \alpha)100\%$ standard normal CI is
\[
(\hat{\theta}_{n} - z_{(1-\alpha/2)}\widehat{\text{SE}}, \quad
\hat{\theta}_{n} - z_{(\alpha/2)}\widehat{\text{SE}}).
\]
\eds{define $z_{(.)}$ and check the signs}
This CI is centered by the point estimator $\hat\theta_n$ and is symmetric.
Its validity relies on whether the distribution of $\hat\theta_n$ is reasonably
well approximated by its asymptotic normal distribution and whether the
bootstrap $\widehat{\text{SE}}$ approximates the true standard error.

\paragraph{Student's $t$ CI}
The procedure for constructing a Student's $t$ CI based on standard bootstrap is
described in \citet[p.158]{efron1993introduction}. \jy{give page or section number when
  citing a book} \mc{addressed}
With block bootstrapping, a $(1 - \alpha)100\%$ Student's $t$ CI is
\[
(\hat{\theta}_{n} - t_{(1-\alpha/2), k - 1}\hat{\text{SE}}, \quad
\hat{\theta}_{n} - t_{(\alpha/2), k -1}\hat{\text{SE}}),
\]
\eds{define $t_{(.)}$ and check the signs}
where $k$ is the number of blocks.
This CI is centered by the point estimator $\hat\theta_n$ and is symmetric.
Like the standard 
normal interval, the Student's $t$ CI is a table-based bootstrap interval
approach. In this case,
its validity relies on whether the distribution of $\hat\theta_n$ is
reasonably well approximation by the $t_{k-1}$ distribution with an
expected value of $\theta$, and whether the bootstrap 
$\widehat{\text{SE}}$ approximates the true standard error.

@@

\jy{Make it flow better and conciser}
The coverage rates of the CIs were used to compare the performance of CIs. Let
$\hat\theta_{L, r}$ and $\hat\theta_{U, r}$ be the lower and upper bound,
respectively, for the confidence interval constructed for each replicate
$r \in \{1, \ldots, R\}$, where $R$ is the
number of replicates.
Then $cov = \#\{\hat\theta_{L, r} < \theta < \hat\theta_{U, r} \}/R$,  
for $r \in \{1, \ldots, R\}$.
If a CI method is valid, then the coverage rate is expected to match the
nominal 
level. Because it is unlikely for the coverage to exactly match the nominal
level,
we can construct a 95\% CI of the coverage
rate (which is an estimate of a proportion with $n = 10000$ \eds{R?}). If the
proportion
0.95 is included in the interval, the block bootstrap method is likely
performing
well. If all values in the interval are below 0.95, the results would suggest 
that the
method either is providing inaccurate estimation, is underestimating the
process' variability, or a combination of both. If all values in the interval
are above .95, the results suggest that the method is overestimating the
process' variability.

@@

Based on Figure~\ref{fig:mu}, 
Student's $t$ CIs appear to have the greatest \eds{largest?} coverage rates with 
respect 
to estimating $\mu$. In fact, for $\phi = -0.4$, Student's $t$ CIs over-cover 
$\mu$ for all sample sizes, whereas other CIs have coverage rates that
hover around the $95\%$ line at this level of $\phi$. For $\phi \geq -0.2$,
$BC_a$ CIs undercover $\mu$ for all $n$ observed. For $\phi = -0.2$, 
Standard, Percentile, %CBP, 
Proposed Method, and $BC$ CIs seem to 
generally cover $\mu$ at the nominal 
rate when $n \geq 400$. Student's $t$ CIs either recover or over-cover $\mu$
for $n \geq 100$. For $\phi = 0$, or when the 
sample is independently normally distributed, $n = 1600$ seems to 
be enough to recover $\mu$ when using Standard, Percentile, 
%CBP, 
Proposed Method, or $BC$ 
CIs. Student's $t$ CIs cover $\mu$ at the nominal rate for $n \geq 100$ at 
$\phi = 0$.
For $\phi = 0.2$, $n > 3200$ seems to be necessary to recover $\mu$
at the nominal level when using
any of the CIs, although some CIs, such as Student's $t$, are much closer to the
$95\%$ line than others, most notably $BC_a$.
For $\phi = 0.4$, a much larger sample size appears to be necessary. Once again,
Student's $t$ CIs are the closest to $95\%$ coverage, and $BC_a$ CIs are the 
furthest.

@@

To summarize, the CI type chosen should be based on the target parameter. 
When estimating $\mu$, the Student's $t$ CI seems to have the highest coverage
overall, although that leads to over-coverage when $\phi$ is -0.4. For a
sample with no temporal dependence, the Student's $t$ CI cover $\mu$ perfectly
at the nominal rate even for small sample sizes. For positive $\phi$, a larger
sample size appears to be necessary regardless of the CI used. In the case
that $\phi$ is unknown and expected to be greater than $0$, $n > 3200$ may be
necessary for block bootstrap interval estimation of $\mu$. If $\phi$ is known,
a lesser sample size may be sufficient. When estimating $\sigma_x$,
$BC_a$ should not be used, and Student's $t$ again has the greatest 
\eds{highest?} coverage
rates. In the case that $\phi$ is unknown and expected to be greater than $0$,
$n > 3200$ may be necessary for block bootstrap interval estimation of
$\sigma_x$. Again, if $\phi$ is known, a lesser sample size may be enough.
When estimating $\phi$, Percentile and $BC_a$ CIs 
deteriorate in coverage, indicating they should not be used in this context. 
Increasing $n$ will not fix the performance of either of these intervals 
because the assumption that the coverage rate will approach the nominal level
as $n$ increases is invalid. Standard and Proposed Method CIs managed to fix the
deterioration problem, and they perform similarly, but Student's $t$ covers
$\phi$ at the nominal rate for much smaller sample sizes. It seems that 
$n = 800$ is large enough to
recover $\phi$ when using the Student's $t$ method.

@@

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{figures/norm_phi_intervals}
  \caption{100 replicate Percentile, Bias-Corrected, and Proposed Method Bootstrap CIs for samples of $n = 800$ for
    an AR(1) process with $\phi = 0.4$. For each replicate, the lower and upper bounds
    of the CIs are displayed, as well as $\hat\theta_n$ and $\bar\theta_n^{(b)}$. }
  \label{fig:npi}
\end{figure}

>>>>

The necessary sample size for bootstrap standard errors to provide valid
uncertainty measures in practice has not been extensively studied. For
independent data and non-block bootstrap, \citet{hesterberg2015teachers} notes
that while percentile-based CIs for the mean \eds{for the mean? differences in means?} \mc{addressed} from 
nonparametric bootstrap
are more accurate than $t$-intervals for larger sample sizes, they are
less accurate for smaller sample sizes. \citet{chernick2009revisiting} finds 
that the best choice of method for estimating
the parameter of a distribution is dependent on the sample size, the number of
bootstrap resamples, and the confidence level. In the context of structural equation
modeling, \citet{nevitt2001performance} find
that a sample size of 200--1000 is usually sufficient for interval estimation
using standard nonparametric bootstrap. For dependent data, the necessary
sample
size for bootstrap is even less clear. In the setting of linear regression with a
focus on estimating the proportion of 
variation attributable to a certain variable, \citet{burch2012nonparametric} finds
that three nonparametric bootstrap methods (Standard, Percentile, and Bias-Corrected
and
Accelerated) approach the coverage of a pivotal quantity as $n$ gets larger for an
underlying normal distribution. However, they find that for other distributions, 
coverage may deteriorate as $n$ increases. 
When attempting to create interval estimations for the Pearson correlation 
coefficient of bivariate normal data, \citet{puth2015variety} find bootstrap 
methods to be less accurate than the Fisher's transformation method even 
for a sample of size $n = 100$ and correlation coefficient equivalent to 0.
Few studies have provided a practical guide on what size is needed
for block bootstrap inference for dependent data.
In a linear regression setting
with
dependent data where the
regression errors are generated from a homoscedastic AR(1) process, 
\citet{goncalves2005bootstrap} find that standard error
estimates from moving 
block bootstrap in small samples may be more accurate than
inference from closed-form asymptotic estimates, but moving block bootstrap 
\jy{did they do block or standard bootstrap?}
\mc{moving block bootstrap}
CIs still do not have sufficient coverage of the target parameter even for sample size
$n = 1024$.
\jy{I added more references to the reference folder that studied coverage of
  bootstrap CIs but not block bootstrap CIs.}

@@

\paragraph{Standard Normal CI}
Assuming that $\hat\theta_n$ is asymptotically normally distributed with
$\theta$ as the mean, we just need an estimate of the standard error to
construct an approximate CI \citep[p.168]{efron1993introduction}.
Let $\widehat{\text{SE}}$ be the empirical standard error of the block bootstrap
copies of $\hat\theta_n^{(b)}$, $b = 1, \ldots, B$. Let $z_{(\alpha)}$ be the critical value for the $100\alpha$th percentile of the standard normal distribution.
A $(1 - \alpha)100\%$ standard normal CI is
\[
(\hat{\theta}_{n} - z_{(1-\alpha/2)}\widehat{\text{SE}}, \quad
\hat{\theta}_{n} - z_{(\alpha/2)}\widehat{\text{SE}}).
\]
\eds{define $z_{(.)}$ and check the signs} \mc{addressed}
This CI is centered by the point estimator $\hat\theta_n$ and is symmetric.
Its validity relies on whether the distribution of $\hat\theta_n$ is reasonably
well approximated by its asymptotic normal distribution and whether the
bootstrap $\widehat{\text{SE}}$ approximates the true standard error.

\paragraph{Student's $t$ CI}
The procedure for constructing a Student's $t$ CI based on standard bootstrap is
described in \citet[p.158]{efron1993introduction}. \jy{give page or section number when
  citing a book} \mc{addressed} Let $t_{(\alpha, k)}$ be the critical value for the $100\alpha$th percentile of the student's $t$ distribution with $k$ degrees of freedom.
With block bootstrapping, a $(1 - \alpha)100\%$ Student's $t$ CI is
\[
(\hat{\theta}_{n} - t_{(1-\alpha/2), k - 1}\hat{\text{SE}}, \quad
\hat{\theta}_{n} - t_{(\alpha/2), k -1}\hat{\text{SE}}),
\]
\eds{define $t_{(.)}$ and check the signs}  \mc{addressed}
where $k$ is the number of blocks.
This CI is centered by the point estimator $\hat\theta_n$ and is symmetric.
Like the standard 
normal interval, the Student's $t$ CI is a table-based bootstrap interval
approach. In this case,
its validity relies on whether the distribution of $\hat\theta_n$ is
reasonably well approximation by the $t_{k-1}$ distribution with an
expected value of $\theta$, and whether the bootstrap 
$\widehat{\text{SE}}$ approximates the true standard error.

@@

\jy{Make it flow better and conciser}
The coverage rates of the CIs were used to compare the performance of CIs. Let
$\hat\theta_{L, r}$ and $\hat\theta_{U, r}$ be the lower and upper bound,
respectively, for the confidence interval constructed for each replicate
$r \in \{1, \ldots, R\}$, where $R$ is the
number of replicates.
Then $cov = \#\{\hat\theta_{L, r} < \theta < \hat\theta_{U, r} \}/R$,  
for $r \in \{1, \ldots, R\}$.
If a CI method is valid, then the coverage rate is expected to match the
nominal 
level. Because it is unlikely for the coverage to exactly match the nominal
level,
we can construct a 95\% CI of the coverage
rate (which is an estimate of a proportion with $R = 10000$ \eds{R?} \mc{addressed}). If the
proportion
0.95 is included in the interval, the block bootstrap method is likely
performing
well. If all values in the interval are below 0.95, the results would suggest 
that the
method either is providing inaccurate estimation, is underestimating the
process' variability, or a combination of both. If all values in the interval
are above .95, the results suggest that the method is overestimating the
process' variability.

@@

Based on Figure~\ref{fig:mu}, 
Student's $t$ CIs appear to have the highest \eds{largest?} \mc{addressed} coverage rates with 
respect 
to estimating $\mu$. In fact, for $\phi = -0.4$, Student's $t$ CIs over-cover 
$\mu$ for all sample sizes, whereas other CIs have coverage rates that
hover around the $95\%$ line at this level of $\phi$. For $\phi \geq -0.2$,
$BC_a$ CIs undercover $\mu$ for all $n$ observed. For $\phi = -0.2$, 
Standard, Percentile, %CBP, 
Proposed Method, and $BC$ CIs seem to 
generally cover $\mu$ at the nominal 
rate when $n \geq 400$. Student's $t$ CIs either recover or over-cover $\mu$
for $n \geq 100$. For $\phi = 0$, or when the 
sample is independently normally distributed, $n = 1600$ seems to 
be enough to recover $\mu$ when using Standard, Percentile, 
%CBP, 
Proposed Method, or $BC$ 
CIs. Student's $t$ CIs cover $\mu$ at the nominal rate for $n \geq 100$ at 
$\phi = 0$.
For $\phi = 0.2$, $n > 3200$ seems to be necessary to recover $\mu$
at the nominal level when using
any of the CIs, although some CIs, such as Student's $t$, are much closer to the
$95\%$ line than others, most notably $BC_a$.
For $\phi = 0.4$, a much larger sample size appears to be necessary. Once again,
Student's $t$ CIs are the closest to $95\%$ coverage, and $BC_a$ CIs are the 
furthest.

@@

To summarize, the CI type chosen should be based on the target parameter. 
When estimating $\mu$, the Student's $t$ CI seems to have the highest coverage
overall, although that leads to over-coverage when $\phi$ is -0.4. For a
sample with no temporal dependence, the Student's $t$ CI cover $\mu$ perfectly
at the nominal rate even for small sample sizes. For positive $\phi$, a larger
sample size appears to be necessary regardless of the CI used. In the case
that $\phi$ is unknown and expected to be greater than $0$, $n > 3200$ may be
necessary for block bootstrap interval estimation of $\mu$. If $\phi$ is known,
a lesser sample size may be sufficient. When estimating $\sigma_x$,
$BC_a$ should not be used, and Student's $t$ again has the highest
\eds{highest?} \mc{addressed} coverage
rates. In the case that $\phi$ is unknown and expected to be greater than $0$,
$n > 3200$ may be necessary for block bootstrap interval estimation of
$\sigma_x$. Again, if $\phi$ is known, a lesser sample size may be enough.
When estimating $\phi$, Percentile and $BC_a$ CIs 
deteriorate in coverage, indicating they should not be used in this context. 
Increasing $n$ will not fix the performance of either of these intervals 
because the assumption that the coverage rate will approach the nominal level
as $n$ increases is invalid. Standard and Proposed Method CIs managed to fix the
deterioration problem, and they perform similarly, but Student's $t$ covers
$\phi$ at the nominal rate for much smaller sample sizes. It seems that 
$n = 800$ is large enough to
recover $\phi$ when using the Student's $t$ method.

@@

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{figures/norm_phi_intervals}
  \caption{100 replicate Percentile, Bias-Corrected, and Proposed Method Bootstrap CIs for samples of $n = 800$ for
    an AR(1) process with $\phi = 0.4$. For each replicate, the lower and upper bounds
    of the CIs are displayed, as well as $\hat\theta_n$ (blue) and $\bar\theta_n^{(b)}$ (red). }
  \label{fig:npi}
\end{figure}
